Data Analysis of Mental Health Trends from 2014 to 2023 in Toronto

### 2024-01-23 outputs/paper/paper.qmd
finish paper 1/23

The visualization in @fig-total and @tbl-sum_occurences_new illustrates a gradual rise in the occurrences of MHA apprehensions from 2014 to 2021, peaking in 2021. Moreover, a slight decrease is observed from 2021 to 2023.

@@

```{r}
#| echo: FALSE
#| message: FALSE
#| label: fig-total
#| fig-cap: Total number of MHA apprehensions in each year

@@

The scatter plot @fig-percentage is used to make it easier to see the trend between the year and the number of MHA apprehensions in different sex. As depecited in @fig-percentage, it can be clearly observed that the number of male MHA apprehensions consistently surpasses that of female MHA apprehensions across the years 2014 to 2023. Futhermore, an observation can be made that in the year of 2020, the male MHA apprehension attained the highest number, while in the year of 2015, the female MHA apprehension attained the highest number across 2014 to 2023. There is also a decrease in the number of MHA apprehensions in both sexes in 2017 compared to 2015 and 2016. It is also worth noting that in the year of 2020, while the number of male MHA apprehensions reaches its peak, the number of female MHA apprehensions is at its lowest. Also, in 2015, the reverse happens: while the number of female MHA apprehensions reaches its peak, the number of male MHA apprehensions is at its lowest. This phenomenon is something worth analyzing in depth, and can be better understood by researching.

```{r}
#| echo: FALSE
#| message: FALSE
#| label: tbl-sum_occurences_sex
#| tbl-cap: Percentage of Male and Female MHA apprehensions in each year

# calculate the percentage of male and female MHA apprehensions in total
percentage <- mental_health_data %>%
  group_by(occurred_year, sex) %>%
  summarise(count = n()) %>%
  group_by(occurred_year) %>%
  mutate(percentage = count / sum(count) * 100)

@@

```{r}
#| echo: FALSE
#| message: FALSE
#| label: fig-percentage
#| fig-cap: Percentage of male and female MHA apprehension percentage from 2014 to 2023
# draw a scatter plot for the percentage of male and female in each year
percentage <- mental_health_data %>%
  group_by(occurred_year, sex) %>%
  summarise(count = n()) %>%
  group_by(occurred_year) %>%
  mutate(percentage = count / sum(count) * 100)

@@

@tbl-sum_occurences and @fig-age_year displays the total occurrences of MHA apprehensions in different age groups in Toronto from 2014 to 2023. It can be observed that the age group "25 to 34" has the highest number of MHA apprehensions. Then the number of MHA apprehensions gradually decreases as the age group becomes older. Age group "18 to 24" and age group "35 to 44" have similar number of MHA apprehensions, which only differs by 333 occurrences.

@@

@@

```{r}
#| echo: FALSE
#| message: FALSE
#| label: fig-age_year
#| fig-cap: Total number of MHA apprenhension in age groups
# Analyze the relationship between MHA apprehension and age groups in total
mental_health_data |> 
  count(age_group) |> 
  ggplot(mapping = aes(x = age_group, y = n)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Age group", y = "Number of MHA apprenhensions")
```

```{r}
#| echo: FALSE
#| message: FALSE
#| label: fig-age_year_each
#| fig-cap: Number of MHA apprenhension in age groups in each year
# Analyze the relationship between MHA apprehension and age groups in each year
mental_health_data |>
  ggplot(mapping = aes(x = occurred_year, fill = age_group)) +
  geom_bar(position = "dodge2") +
  labs(x = "Year", y = "Number of MHA apprenhension", fill = "Age group") +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = unique(percentage$occurred_year))
```

@@

>>>>

The visualization in @fig-total and @tbl-sum_occurences_new illustrates a gradual rise in the occurrences of MHA apprehensions from 2014 to 2021, peaking in 2021 (13219 occurrences). Moreover, a slight decrease is observed from 2021 to 2023.

@@

```{r out.width="65%", fig.align='center'}
#| echo: FALSE
#| message: FALSE
#| label: fig-total
#| fig-cap: Total number of MHA apprehensions in each year

@@

The scatter plot @fig-percentage is used to make it easier to see the trend between the year and the number of MHA apprehensions in different sex. As depecited in @fig-percentage, it can be clearly observed that the number of male MHA apprehensions consistently surpasses that of female MHA apprehensions across the years 2014 to 2023. Futhermore, an observation can be made that in the year of 2020, the male MHA apprehension attained the highest percentage (58.55%), while in the year of 2015, the female MHA apprehension attained the highest percentage (44.26%) across 2014 to 2023. There is also a decrease in the number of MHA apprehensions in both sexes in 2017 compared to 2015 and 2016. It is also worth noting that in the year of 2020, while the number of male MHA apprehensions reaches its peak, the number of female MHA apprehensions is at its lowest (differs by 17.1%). Also, in 2015, the reverse happens: while the number of female MHA apprehensions reaches its peak, the number of male MHA apprehensions is at its lowest (differs by 11.48%). This phenomenon is something worth analyzing in depth, and can be better understood by researching.

Overall, it can be concluded that males have a higher number of MHA apprehensions that females.

```{r tbl.}
#| echo: FALSE
#| message: FALSE
#| label: tbl-sum_occurences_sex
#| tbl-cap: Percentage of Male and Female MHA apprehensions in each year

# calculate the percentage of male and female MHA apprehensions in total
percentage <- mental_health_data %>%
  group_by(occurred_year, sex) %>%
  summarise(count = n()) %>%
  group_by(occurred_year) %>%
  mutate(percentage = format(round((count / sum(count) * 100),2), nsmall = 2))

@@

```{r out.width="70%", fig.align='center'}
#| echo: FALSE
#| message: FALSE
#| label: fig-percentage
#| fig-cap: Percentage of male and female MHA apprehension percentage from 2014 to 2023
# draw a scatter plot for the percentage of male and female in each year

@@

@tbl-sum_occurences and @fig-age_year displays the total occurrences of MHA apprehensions in different age groups in Toronto from 2014 to 2023. It can be observed that the age group "25 to 34" has the highest number of MHA apprehensions (30186 occurrences). Then the number of MHA apprehensions gradually decreases as the age group becomes older. Age group "18 to 24" and age group "35 to 44" have similar number of MHA apprehensions, which only differs by 333 occurrences.

@@

Overall, it can be implied that individuals in age group "25 to 34" have the highest number of MHA apprehensions compared to other age groups.

@@

```{r out.width="70%", fig.align='center'}
#| echo: FALSE
#| message: FALSE
#| label: fig-age_year
#| fig-cap: Total number of MHA apprenhension in age groups
# Analyze the relationship between MHA apprehension and age groups in total
mental_health_data |> 
  count(age_group) |> 
  ggplot(mapping = aes(x = age_group, y = n)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Age group", y = "Number of MHA apprenhensions")
```

```{r out.width="70%", fig.align='center'}
#| echo: FALSE
#| message: FALSE
#| label: fig-age_year_each
#| fig-cap: Number of MHA apprenhension in age groups in each year
# Analyze the relationship between MHA apprehension and age groups in each year
mental_health_data |>
  ggplot(mapping = aes(x = occurred_year, fill = age_group)) +
  geom_bar(position = "dodge2") +
  labs(x = "Year", y = "Number of MHA apprenhension", fill = "Age group") +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = unique(percentage$occurred_year))
```

@@

Overall, a conclusion can be reached that individuals living in "Apartment" have the highest occurrences of MHA apprehension.
